---
title: "Data607 - Project 1"
author: "Juan Falck"
date: "Feb 20, 2022"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

## Chess Tournament

We will attempt to complete this project by using ony **REGEX** (Regular Expressions) to extract all features necessary. I am sure there are many other ways to tackle this project, but since I've never used REGEX before, I decided to give it a try.

### Initialization

We won't use many packages, just the basic ones in **TIDYVERSE**

```{r}
rm(list=ls())
library(tidyverse)
```

### Load the chess.txt file

We will load the TXT file to manipulate it with R using **regular expressions** only. We will skip the first 4 lines, since they don't have any information we will use in this project.

```{r}
my_chess <- read_lines("chess.txt", skip=4)
head(my_chess)
```


### REGEX Matching and extracting features

For each feature we will use the same procedure:  We will define a pattern for REGEX.  We will use **str_match_all** to extract matching features.  We will remove any unwanted artifacts, like empty rows or columns. Trim any white-space and convert to numeric if necessary

#### Extract player's NAMES


```{r}
pattern_names <- "(?<=\\| )[A-Za-z -]{6,}(?=\\|)"
names <- str_match_all(my_chess,pattern_names)

# The match added blank lines so I remove every 2,3 rows and 
# remove all leading and trailing white space
names2 <- str_trim(names[seq(1, length(names), 3)],
                   side = c("both","left","right"))
head(names2)
```


#### Extract player's STATES


```{r}
# Pattern for STATE
pattern_states <- "[A-Z]{2}(?=\\s\\|)"
states <- str_match_all(my_chess,pattern_states)
states2 <- str_trim(states[seq(2, length(states), 3)],
                   side = c("both","left","right"))
head(states2)
```


#### Extract player's POINTS


```{r}
pattern_points <- "\\d\\.\\d"
points <- str_match_all(my_chess,pattern_points)
points2 <- str_trim(points[seq(1, length(points), 3)],
                    side = c("both","left","right"))
head(points2)
```


#### Extract PLAYER's RATING


```{r}
# Pattern Rating
pattern_rating <- "((?<=R: )|(?<=R:  ))\\d{3,4}"
rating <- str_match_all(my_chess,pattern_rating)
#Remove unneeded rows
rating2 <- rating[seq(2, length(rating), 3)]
#Remove unneeded columns
rating_temp <- lapply(rating2, function(x) x[,-2])
# Trim whitespace
rating2 <- str_trim(rating_temp,side = c("both","left","right"))
# Convert from character to numeric
rating2 <- unlist(lapply(rating2,as.numeric), recursive = FALSE)
head(rating2)
```


#### Extract PLAYERS PLAYED


```{r}
# This line extract every 3rd row
my_chess_games <- my_chess[seq(1, length(my_chess), 3)]

# This pattern extracts the number that followed a D, W or L"
pattern_games <- "(((?<=W )|(?<=L ))|(?<=D ))\\s{0,3}\\d{0,2}(?=\\|)"

# Results are stored in this list of lists
players_played <- str_match_all(my_chess_games,pattern_games)

# Lets remove 2 and 3 column, since I won't use them
players2 <- lapply(players_played, function(x) x[,-c(2:3)])

# Need to trim for whitespace
players2 <- lapply(players2,str_trim)
head(players2)
```


#### Calculate AVERAGE RATING OF PLAYERS PLAYED 

This one was tricky. I have all data elements I need to calculate the means of ratings, I was looking for a way to do it without a for loop. Unfortunately don't know enough R to figure out a sleak way of applying the MEAN to the vectors I had and store the results as a simple vector. **LAPPLY** would generate the **AVERAGES** I need, but the return was a very complex multi-dimensional LIST of LISTS of LISTS.  So I gave up and did the FOR LOOP.


```{r}
index_players <- lapply(players2,as.numeric)
rows_players <- length(index_players)
player_opponents <- vector(mode = "list", length = rows_players)

for (row in 1:rows_players) {
  for (col in 1:length(index_players[[row]])) {
    player_opponents[[row]][col] <- rating2[[index_players[[row]][col]]]
  }
}

# We need to take mean and then simplify list of vectors into a single vector
avg_players <- round(unlist(lapply(player_opponents,mean),recursive=FALSE))

head(avg_players)
```


### Create CSV file

Now we have all the information we need. We will merge it into comma separated file

```{r}
output <- cbind(names2,states2,points2,rating2,avg_players)
write.csv(output,"project1.txt", row.names=FALSE)

```

### Let's check the CSV to make all is OK

Final thing to do is to load the file and inspect it to see if results are exactly what we wanted.

```{r}
check_csv <- read_csv("project1.txt")
head(check_csv)

```


**All good!**

# THANK YOU!

