---
title: "Data607 - Assignment Week 3"
author: "Juan Falck"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

Let's first delete all variables and load our tidyverse library.

```{r load-packages, message=FALSE}
rm(list=ls()) 
library(tidyverse)
```

## Exercise 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

**ANSWER** 
We will load directly from github using the alternatve site which works for read_csv. I will use regex and the str_detect function in Stringr to match any major whose description either has the word "DATA" or the word "STATISTICS.

Since all descriptions were already CAPITALIZED, I didn't have to transform any of the columns or adapt the regex to account for it.

```{r code-chunk-label1}
# Insert code for Exercise 1 here
college_majors <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

# I applied a simple FILTER to get requested output
data_majors <- college_majors %>%
  filter(str_detect(Major, "DATA|STATISTICS"))

data_majors
```

## Exercise 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


**ANSWER**
This one was more complicated than I thought.  The complication is that I was not exactly certain on the format of the output.

The complication was how to threat the quotation chartacters " since the output has in itself quotations.  

So my interpretation was to provide back a string that has quote characters within the string itself, or something like 
**Answer_String <- "c("word1","word2", "etc")"**.  

But this means that the answer string must include escapes and should be something like **Answer_String <- "c(\"word1\",\"word2\",\"etc\")"**. 

This way if I write this code below the output will look like what I think the output should look like:

```{r code-chunk-label12}
Answer_String <- "c(\"word1\",\"word2\",\"etc\")"
 writeLines(Answer_String)
```

So I tried this several ways using variations of REGEX. Let's first load the text from a file I created. The approach I took in this problems is to read the **whole text in a single variable**.

```{r code-chunk-label2}
my_file <- read_file("words.txt")
my_file
```

Then for approach #1 I decided to use REGEX to match the whole word INCLUDING THE QUOTATION characters

```{r code-chunk-label3}
# Insert code for Exercise 1 here
pattern <- '["“][A-Za-z ]{2,}["”]'
str_view_all(my_file,pattern)
data1 <- str_match_all(my_file,pattern)
```

The match was as designed, and the result stored in a LIST.  

Next step was to merge the words in quotes.


```{r code-chunk-label4}
data1 <- paste(data1,collapse=",")
writeLines((data1))
```

The result looks VERY CLOSE to desired answer, but not exactly as I desired since it added quotation characters around each word. So instead of ending with "word1", I ended with ""word1"". Not what I wanted.

So I went the route to match the words inside quotation marks, but without the quotation characters themselves, just the words inside.  I used again REGEX to isolate words.

```{r code-chunk-label5}
pattern <- '(?<=\")[A-Za-z ]{3,}(?=\")'
str_view_all(my_file,pattern)
data2 <- str_match_all(my_file,pattern)
```
 
I almost got what I wanted. Unfortunately it also added some blank spaces between words in quotes. After some experimenting coudlnt' find a REGEX that would I isolate just rhe words within qoutes. 
 
I decided just to go ahead remove all empty strings added to the list. Not pretty but would get me there


```{r code-chunk-label6}
data3 <- data2[[1]][str_count(data2[[1]],pattern=" ") < 2]
data3
```

So that gives me what I want, a list of the just words in the file, without quotes around them.

My next step was to merge the words in the list with "," as separator.


```{r code-chunk-label8}
data4<- paste(data3,collapse = '\",\"')
writeLines(data4)
```

So this ALMOST completely did the trick. I just need to add **"c("** at the beginning and **")** at the end and PRESTO.


```{r code-chunk-label9}
final_string <- paste0('c(\"',data4,'\")')
writeLines(final_string)
```

**Exacty the output requested in the exercise DONE!**


## Exercise 3

Describe, in words, what these expressions will match:

(.)\1\1

"(.)(.)\\2\\1"

(..)\1

"(.).\\1.\\1"

"(.)(.)(.).*\\3\\2\\1"


**ANSWER**

Let's start by defining a test vector we can use to test our REGEX.

```{r code-chunk-label10}
test <- c("bell pepper", "aaairlines", "bilberry", "blackberry", "blood orange", 
  "blueberry", "cantaloupe", "chili pepper", "church","cloudberry", 
  "elderberry", "lime", "rhbbbf", "lychee", "mulberry", "olive",
  "salal berry", "1234ab345 ba343", "1234ab345ba343","alibaba", "bereb", 
  "bbgun", "asdbbb", "balacab", "balacaba", "1234568965443", "erberberbee")

```


Now lets look and test what each REGEX do

```{r code-chunk-label10b}

#Searches the whole string and would match 3 continuous of same character

str_view_all(test,"(.)\\1\\1")

#Searches the whole string and looks for 4 characters, where
#the last pair of two is in reverse order than the first pair of two

str_view_all(test,"(.)(.)\\2\\1")

#Searches and look for 4 consecutive characters where the first
#pair of characters is repeated once

str_view_all(test,"(..)\\1")

#Looks for a pattern of same character repeated 3 times with a character
#in between

str_view_all(test,"(.).\\1.\\1")

```


## Exercise 4

Construct regular expressions to match words that:

- Start and end with the same character.
- Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
- Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


**ANSWER**
Let's look at the REGEX I came up with for each of the cases.

```{r code-chunk-label11}
# Start and end with the same character.

str_view_all(test,"^(.).*\\1$")

# Contain a repeated pair of letters (e.g. "church" contains 
# "ch" repeated twice.)

str_view_all(test,"(..).*\\1")

# or

str_view_all(test, "([A-Za-z][A-Za-z]).*\\1")


# Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)   

str_view_all(test,"([a-z]|[A-Z]).*\\1.*\\1")

```


**Thank you!**


...

