---
title: "DATA 607 Assignment 1"
author: "Juan Falck"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(reshape)
```

# Introduction
For this assignment I used the dataset contained in the article "How Baby Boomers get high". The article link is here [link](https://fivethirtyeight.com/features/how-baby-boomers-get-high/)
The artcile analyzes drug usage (including alcohol) by different age-groups.  The article focuses on baby-boomers defined in this article as people ages 50-64. General findings were that boomers consume drugs in lower percentages than younger generations, but they were consuming it at higher rates than their parents.  


# Sub-setting of data
Just for fun for this excercise I will aim at comparing my age group (50-64) to my daughter's age group (21 years old). Let's see what I find.


## Data Manipulation Steps

Load the data directly from site. In this case we will use the drug use by age-group.

```{r p1}
drug_use <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv")
```


Now we will remove the columns we don't need. We will focus of usage numbers

```{r p2}
drug_use <- drug_use %>% select(-contains("frequency"))
```

Now we will focus in my age group 50-64

```{r p3}
drug_use <- drug_use %>% filter(age=="50-64" | age=="21")
```


Lets remove some columns we don't need

```{r p4}
drug_use <- drug_use %>% select(-c(n))
```

Lets transpose for ease of view and save it to another data frame

```{r p5}
drug_use2 <- as_tibble(cbind(drug = names(drug_use), t(drug_use)))
```

Remove first row (age)

```{r p5B}
drug_use2 <-  filter(drug_use2,drug != 'age')
```


Convert to numeric

```{r p6}
drug_use2$V2 <- as.numeric(drug_use2$V2)
drug_use2$V3 <- as.numeric(drug_use2$V3)
```

Lets sort it

```{r p7}
drug_use2 <- arrange(drug_use2,desc(V2)) %>%
  dplyr::rename(Group_21_YO = V2, Group_50_64_YO =V3)
drug_use2
```

Graph them, but first convert to long format from wide format

```{r p8}
data_long <- gather(drug_use2, age_group, drug_use_pct, Group_21_YO:Group_50_64_YO, factor_key=TRUE)

ggplot(data_long, # the data that I am using
       aes(x = drug, # 'aesthetic' includes x
           y = drug_use_pct, fill=age_group)) + # and y
  geom_bar(position="dodge", stat = "identity") + # use ACTUAL y for bar height
  coord_flip()

```

# Conclusions
## From the article
I would recommend repeat the study since it has been 10 year since this study came out.  One of findings was the trend of baby boomers cosuming more drugs than their parents.  After 10 yeasr we could repeat and see if this new cohort of 50-64 has continued the increasing trend or not.

## From my comparison to my daughter's age group
It was evident my age group consume much less drugs than 21 yo.  In fact besides alcohol and marijuana, my age group has little consumption of other drugs vs 21 yo. The article mentioned that consumption in younger people is more about getting high vs older people who if they use it is to cope with stress and other underlying issues.

...

