---
title: "Data 607 Assignment Week 5"
author: "Juan Falck"
date: "2/28/2022"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"
  )   # slightly smaller font for code
```

## Tidying and Transforming  Vaccination  Data


### Initialization

I will use the package "readxl" to read directly the .XLS file from my drive.

```{r}
library(tidyverse)
library(readxl)
library(httr)
rm(list=ls())
```

### Read in the .XLS directly from github folder 

I will use the **GET** function from **httr** package to read the file from Github and save it on a temporary folder on my windows folder. 
Then I will use the **read_excel** function from **readxl** package to read directly the .XLS file from the temporary folder

```{r }

url1 <- "https://raw.githubusercontent.com/acatlin/data/master/israeli_vaccination_data_analysis_start.xlsx"

GET(url1, write_disk(tf <- tempfile(fileext = ".xlsx")))

my_excel <- read_excel(tf)
my_excel
```

### Some tidying up

The tibble is absolutely messy.  I will delete all rows and columns I dont need.

So I decided I only need **rows 2,3,4,5 and the first 4 columns**. Everyrhing else I will delete.


```{r}
#Delete Rows I don't need.
my_excel2 <- my_excel %>% slice(-c(1,6:15))
my_excel2
```

It looks much better now. We don't need last column

```{r}
my_excel2 <- my_excel2 %>%
  select(-Efficacy)
my_excel2
```


I don't like the column names so I will change them to:
age_group, pop_n_vax, pop_f_vax, severe_nvax, severe_vax


```{r}
my_excel3 <- my_excel2 %>% 
  rename(
    age_group = Age,
    pop_n_vaxed = `Population %`,
    pop_f_vaxed = ...3,
    severe_nvax = `Severe Cases`,
    severe_vax = ...5)
str(my_excel3)
```

Population numbers are now characters so I will convert them to numbers

```{r}
cols.num <- c("pop_n_vaxed","pop_f_vaxed","severe_nvax","severe_vax")

my_excel3[,cols.num] <- sapply(my_excel3[,cols.num],as.numeric)
str(my_excel3)
```

### Write to CSV

One last step before I write to csv is to move the percentage in the two rows to the respective column

```{r}
#First lets add two new empty columns
my_excel4 <- my_excel3 %>%
  add_column(pct_f_vax = NA,
             pct_n_vax = NA)
#now the pcts to the proper row and column
my_excel4[1,6]<- my_excel4[2,3]
my_excel4[1,7]<- my_excel4[2,2]
my_excel4[3,6]<- my_excel4[4,3]
my_excel4[3,7]<- my_excel4[4,2]

# Finally delete the two rows you don't need
my_excel4 <- my_excel4 %>% slice(-c(2,4))
my_excel4
```

Looks good and ready to send to csv

```{r}
write.csv(my_excel4,"covid.csv", row.names = FALSE)

```

Let's read it back to see it all is ok
```{r}
my_csv <- read_csv("covid.csv")
my_csv
```

**PERFECT!**


### Question 1

Do I have enough information to calculate the total population?

**ANSWER**

Yes. One important thing to note is that the total population in the table is not 100%.  I am assuming there is a category "vaxed but not fully" which is not in the table.  So in order to get full population we would need to do some basic algebra.

```{r}
my_csv2 <- my_csv %>%
  mutate(pct_s_vax = 1-(pct_f_vax+pct_n_vax),
         pop_s_vax = (1-(pct_f_vax+pct_n_vax)) * pop_f_vaxed/pct_f_vax)

# Total population single vax by Age group
my_csv2[,c("age_group","pop_s_vax")]

#Total pct of population single vax by age group
my_csv2[,c("age_group","pct_s_vax")]

```

Now we have all we need to output full population

```{r}
my_csv2 %>%
  summarise(Age_group = age_group,
            total = pop_n_vaxed+pop_f_vaxed+pop_s_vax,
            pct = pct_f_vax+pct_n_vax+pct_s_vax
            )
```


### Question 2

Calculate the efficacy ratio

**ANSWER**

Now that we have our tibble lest add a column for efficacy.
using the formula **1- RiskRatio** where Risk Ratio is equal Risk_vaxed / Risk_unvaxed


```{r}
my_csv3 <- my_csv2 %>%
 mutate(efficacy = 1-(severe_vax/severe_nvax))

my_csv3$efficacy
```

The results seem counter intuitive.  For the age group 50 below, the results are directionally what we expected that the efficacy of the vaccine was 74% vs un-vaccinated people.

Yet the efficacy for the poulation over 50% seems to be in the other direction where the rate for severe diseases was higher for vaccinate people vs un-vaccinated.  There could be many reasons for this counter intuitive result for example underlying conditions which could easier trigger a severe case from what normally would've been mild. 

### Question 3

Can we compare total rates for vaccinated vs unvaccinated?

**ANSWER**

Yes, but we need to do some algebra again. 


```{r}
# we need to calculate first the total number of severe cases
my_csv3 %>%
  mutate(total_severe_vax = pop_f_vaxed/100000*severe_vax,
         total_severe_nvax = pop_n_vaxed/100000*severe_nvax) %>%
  summarise(rate_vaxed = total_severe_vax / (sum(pop_f_vaxed) / 100000),
            rate_nvaxed = total_severe_nvax / (sum(pop_n_vaxed) / 100000)) %>%
  summarise_all(sum)
```


We can see now that the blended rate for the population tells us that severe cases per 100K is **117 for vaccinated people vs 61.3 for unvaccinated**.  

This is very counter-intuitive result but it is what the numbers are telling me.


**Thanks!**
