---
title: "Assignment 6 - XML/JSON/HTML"
author: "Juan Falck"
date: "March 17,2022"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",
  warning = FALSE, message = FALSE)   # slightly smaller font for code
```

## Introduction
I have placed three files in the the three formats (XML, JSON, HTML) in a AWS S3 folder so they can be accesed directly from the internet.  Links are below:

https://cuny-msds.s3.amazonaws.com/books.xml

https://cuny-msds.s3.amazonaws.com/books.json

https://cuny-msds.s3.amazonaws.com/books.html

We will ready the files, parse them and ensure they are within an R dataframe or tibble.


## XML to R

Let's load the libraries we will use.

```{r}
rm(list=ls())
library(tidyverse)
library(XML)
library(xml2)
```

Let's load the files directly from our AWS S3 folders.

```{r}
url <- "https://cuny-msds.s3.amazonaws.com/books.xml"
data_xml <- read_xml(url)
books_xml <- xmlParse(data_xml)
df_books_xml <- xmlToDataFrame(nodes = getNodeSet(books_xml, "//book"))
```

Let's check the structure and look at the data

```{r}
str(df_books_xml)
```

Some numerical values were imported as character. So let's convert them to numerical

```{r}
df_books_xml$year <- as.numeric(as.character(df_books_xml$year))
df_books_xml$amazon_price <- as.numeric(as.character(df_books_xml$amazon_price))
str(df_books_xml)
df_books_xml
```

Now it looks all good. Let's move on!

## From json to R

Let's load libraries

```{r}
library("jsonlite")
library("rjson")
```

Let's load the data into R.

```{r}
url <- "https://cuny-msds.s3.amazonaws.com/books.json" 
data_json <- jsonlite::fromJSON(url)
df_json <- as.data.frame(data_json)
```

Let's check the structure

```{r}
str(df_json)
df_json
```

All looks good! let's continue.

## HTML to R

We will use the rvest library to load HTML into R.

```{r}
library(rvest)
```

Load the data from the S3 folder in our AWS account

```{r}
url <- "https://cuny-msds.s3.amazonaws.com/books.html"
data_html <- read_html(url)
table_html <- html_table(data_html, fill=TRUE)[[1]]
```


Let's check the structure and look at the table

```{r}
str(table_html)
table_html
```

All looks good, we are done!

```{r}
```


## Thank you!
